---
title: "Exercise 3"
output: html_document
date: '2022-05-06'
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Task 1: Import Data and Segmentation

```{r message = FALSE}
library(readr)        # to import tabular data (e.g. csv)
library(ggplot2)      # to visualize data
library(dplyr)        # to manipulate (tabular) data
library(tidyr)
caro <- read_delim("caro60.csv",",")
caro
ggplot(caro, aes(E,N))+
  geom_point()+
  geom_path()+
  labs(title= "Caros Trajectory")
```

Guidelines for Segmentation of trajectories: 1. Specify a temporal
windows v for in which to measure Euclidean distances -\>The sampling
rate is 1min and the temporal window v should be 6min = +/- 3min 2.
Measure the distance from every point to every other point within this
temporal window (v) 3. Remove "static points": These are points where
the average distance is less than a given threshold. This segments the
trajectory into subtrajectories -\> Distances under the mean distance
value are considered static in this example 4. Now remove short
subtrajectories: These are trajectories with a short duration (whereas
"short" is tbd)

```{r message = FALSE}
## 2. Measure distance between points ##
caro <- caro %>%
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2+(lag(N,3)-N)^2),   # distance to pos -3min
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),   # distance to pos -2 min
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),   # distance to pos -1 min
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), # distance to pos +1 min
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2), # distance to pos +2 min
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)  # distance to pos +3 min
  )

## Calculate mean distance ##
caro <- caro %>%
  rowwise() %>%
  mutate(
    stepMean = mean(c(nMinus2, nMinus1,nPlus1,nPlus2))
  ) %>%
  ungroup() 

```

Task 2: Specify and apply threshold d

```{r message = FALSE}

## 3. Remove static points ##
caro <- caro %>%                            
  ungroup() %>%
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE)) #Boolean for < mean distance in a new colum


caro_filter <- caro %>%
  filter(!static)         #filters for row with static != TRUE (distances higher than mean distance)

caro_filter%>%
  ggplot(aes(E, N))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")+
  labs(title = "Caros segmented Trajectory")
```

Task 3: Visualize segmented trajectories

```{r message = FALSE}
ggplot(caro, aes(E,N, color = static))  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  theme(legend.position = "bottom")+
  labs(title = "Caros static and non-static Trajectory")
```

Task 4: Segment-based analysis Complete the guidelines, short
trajectories must be removed now. For that we need a function to greate
a unique ID for each sub-trajectory.

```{r message = FALSE}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
}

```

Apply function to caro and visualize the sub-trajectories(Don't forget
to filter the Data by static to exclude non moving segments):

```{r message = FALSE}
caro <- caro %>%
  mutate(segment_id = rle_id(static))

caro_filter_2 <- caro %>%
  filter(!static)

ggplot(caro_filter_2, aes(E,N, color = segment_id), show.legend = FALSE)  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  labs(title = "Caros sub-trajectories")+
  theme(legend.position = "none")

```

Compare output with the output when only segments with duration \>=5min
are shown (Long segments):

```{r message = FALSE}
caro_filter_3 <- caro_filter_2 %>% 
  group_by(segment_id) %>% 
  mutate(duration = as.integer(difftime(max(DatetimeUTC),min(DatetimeUTC), units = "mins"))) %>%
  filter(duration >= 5)

ggplot(caro_filter_3, aes(E,N, color = segment_id), show.legend = FALSE)  +
  geom_path() +
  geom_point() +
  coord_fixed() +
  labs(title = "Caros long sub-trajectories")+
  theme(legend.position = "none")

```

Task 5: Similarity measures

```{r message= FALSE}
## Import Data ##
ped <- read_delim("pedestrian.csv",",")
ped

##Visualization of trajectories ##
ped<- ped%>%
  group_by(TrajID)
ped

ped$TrajID<- as.character(ped$TrajID)
ggplot(ped, aes(E, N))+
  geom_path() + 
  geom_point(aes(color=TrajID)) + 
  scale_color_manual(values = c("1" = "brown2", "2" = "coral", "3"="khaki2", "4"= "light green", "5"= "turquoise", "6"= " dark orchid2"))+
  facet_wrap(~TrajID, nrow = 2, labeller = label_both)+
  labs( title = "Visual comparison of  6 trajectories 01.03.2015",
       subtitle = "Each subplot highlights a unique pedestrian trajectories")
```

The trajectories are very similar, but TrajID 4 shows a slightly
different pattern and there is a clear outlier at 5. The trajectories
also differ in terms of speed. This can be seen by the different density
of the points (since the sampling rate is the same for all of them).

Task 6: Install Library "SimilarityMeasures" & "tidyr"

```{r message=FALSE}
library(SimilarityMeasures)
help(package = "SimilarityMeasures")
```
Calculate similarity Compare trajectory 1 to trajectories 2-6
with different methodes.
```{r message= FALSE}
## Create matrix for each trajectory ##

trj_1<-ped%>%
  filter(TrajID == 1)%>%
  ungroup()%>%
  select(E,N)%>%
  as.matrix()

trj_2 <- ped %>%
  filter(TrajID == 2) %>%
  ungroup()%>%
  select(E, N) %>%
  as.matrix()

trj_3 <- ped %>%
  filter(TrajID == 3) %>%
  ungroup()%>%
  select(E, N) %>%
  as.matrix()

trj_4 <- ped %>%
  filter(TrajID == 4) %>%
  ungroup()%>%
  select(E, N) %>%
  as.matrix()

trj_5 <- ped %>%
  filter(TrajID == 5) %>%
  ungroup()%>%
  select(E, N) %>%
  as.matrix()

trj_6 <- ped %>%
  filter(TrajID == 6) %>%
  ungroup()%>%
  select(E, N) %>%
  as.matrix()

DTW<- c(DTW(trj_1, trj_2),
       DTW(trj_1, trj_3),
       DTW(trj_1, trj_4),
       DTW(trj_1, trj_5),
       DTW(trj_1, trj_6))
DTW

EditDist<- c(EditDist(trj_1, trj_2),
       EditDist(trj_1, trj_3),
       EditDist(trj_1, trj_4),
       EditDist(trj_1, trj_5),
       EditDist(trj_1, trj_6))
EditDist

Frechet<-c(Frechet(trj_1, trj_2),
       Frechet(trj_1, trj_3),
       Frechet(trj_1, trj_4),
       Frechet(trj_1, trj_5),
       Frechet(trj_1, trj_6))
Frechet

LCSS<-c(LCSS(trj_1, trj_2,pointSpacing = 5 ,pointDistance = 3, errorMarg =0.7),
      LCSS(trj_1, trj_3,pointSpacing = 5 ,pointDistance = 3, errorMarg = 0.7),
      LCSS(trj_1, trj_4,pointSpacing = 5 ,pointDistance = 3, errorMarg = 0.7),
      LCSS(trj_1, trj_5,pointSpacing = 5 ,pointDistance = 3, errorMarg = 0.7),
      LCSS(trj_1, trj_6,pointSpacing = 5 ,pointDistance = 3, errorMarg = 0.7))
LCSS

## Combine data as data frame ##
TrajID<-c("2","3","4","5","6")
DTW<-as.integer(DTW)
Frechet<-as.integer(Frechet)
Similarity<- cbind(TrajID,DTW,EditDist,Frechet,LCSS)%>%
  as.data.frame()%>%
  pivot_longer(cols = c(DTW, EditDist, Frechet, LCSS))%>%
  group_by(name)
```

Visualisation of results.
```{r message=FALSE}
ggplot(Similarity, aes(x = TrajID, y = value, fill= TrajID)) +
  geom_bar(stat = "identity")+
  facet_wrap(~name, nrow = 2,
             scales = "free_y")+
  labs(title = "Computed similarities using different measures \nbetween trajectory 1 to all other trajectories")
```
With the the method Edit Distance, every change/difference of the trajectories comes with a cost. A low value means they are very similar. The vlaue for LCSS represents a similarity score, here high values mean the trajectories are mor similar. DTW considers the sum of all squared distances. And the Frechet does the contrary, what is also nicely shown in the graph. However, I have tried various parameter settings and with every change the resluts are very different. Therefore, in a real project, one should consider the data propertiers and effects to find proper settings.

```{r message=FALSE}
